---
title: "Getting Started with Interactive Plotting in R"
author: Jane Carlen, UC Davis Data Science Initiative
date: March 1, 2019
output: 
  ioslides_presentation:
    incremental: false
---

## Goals

Present logic and tools for interactively plotting *static* data. 

In two weeks we'll discuss Shiny, where data can be *dynamic*.

- When to use interactivity
- Work through an examples with
    - Plotly  
    - Crosstalk  
- Other packages, RBokeh, Highcharter  

<!--look more into leaflet-->
<!-- which ones have ability to write your own js to add to event handlers-->


## Why Interactive Plots?

> - ~~They look cool~~
> - Exploratory, especially valuable EDA, your own or the users.
    + Plots you publish have an argument, focus reader attention.
> - Browser-based, easy to share, few compatibility issues.
> - Expose the data
> - Increase the amount of information you can share cleanly
> - Brushing and linking
    + Draw connections between data
    + Uncover relationships

<!-- Best practices for interactivity. One slide? Sprinkled throughout? -->

## Example - Pizza to the Polls

[Pizza to the polls](https://polls.pizza/) - Delivered over 10,000 pizzas to 41 states    
Data from 2018 midterm elections

<!-- I used google map API to get lat/long, standardized variable names, removed some unnecessary fields,
("X", "Other.Info", "Contact.Info", "Slack.Thread", "Pizza.phone.number", "Slack.ID", "Twitter.Thread", "Tweet.ID"),
and removed 14 missing-address rows from the original data -->

```{r setup, echo = F, cache = T, warning = F, messsage = F}
library(dplyr)
library(ggplot2)
library(plotly)
library(crosstalk)
```


```{r load, cache = F}
setwd("~/Documents/DSI/intro_to_interactive")
pizza = read.csv("pizza.csv")
names(pizza); dim(pizza)
```

## Pizza data {.smaller}

```{r}
head(pizza, 4)
```

## How to visualize this data? {}

Questions we have:
- Where were pizzas delivered? 
- When were pizzas delivered? Relative to poll-closing times?
- What describes places where people wait on lines?

## Example - Static Plot

Here is a static plot built in ggplot. It looks OK, but what is it missing?

## Example - Interactive Plot

<!--Pizza plot or other example -->

## Example - Linked Interactive Plot

<!--Pizza plot or other example -->

```{r election_data, eval = F}

# NOT the elections package on cran - https://github.com/MEDSL/elections/blob/master/README.md
# make sure devtools is updated
if (!require('devtools', quietly = TRUE)) install.packages('devtools')
devtools::install_github('MEDSL/elections') 

# The package makes available the following datasets:

# presidential_precincts_2016
# senate_precincts_2016
# house_precincts_2016
# state_precincts_2016
# local_precincts_2016

data("presidential_precincts_2016")
head(presidential_precincts_2016)

pres_by_state_returns_2016 = presidential_precincts_2016 %>% group_by(state_postal, party) %>%
  summarize(part_votes = sum(votes))

https://github.com/nvkelso/election-geodata

```


<!--  Closeness of election outcomes --> 
<!-- filters on pizzas to show --> ?

## Live Example

<!-- Now we code an example ourselves (leave presentation). Stay with pizza example and data?-->

## Review - Advantages of Plotly

## Review - Limitations of Plotly

## Related Packages - Bokeh

## Related Packages - Highcharter

## Related Packages - rCharts



## Part II - Under the hood (Duncan)

- What is plotly actually doing -- relationship between R, html, javascript
- What kinds of backends (html, server, server running R) can be how and how do they work?
