---
title: "Decision Tree Models of Burn Data"
output:
  html_document: default
  pdf_document: default
---

This is proof-of-concept work to show that we can improve upon existing criteria to detect bloodstream infection in burn patients based on vital variables. 

This work has four parts:

1. Compare best models derived from Greenhalgh and Sirs variables (separately) to current Greenhalgh and SIRS criteria. The models may be simple cutoffs, logistic regression, and/or decision trees. 

2. Same as above but combining Greenhalgh and SIRS variables.

3. (If time) Consider demographic variables, e.g. age, gender and interactions, but still limited to single time point data.

4. (If time) Consider derived variables like cumulative days above a threshold. This cant be implemented currently, but could help argue for expanded capabilities with medical records. 

```{r setup, cache = T, results = "hide", fig.keep = "none", warning = F, message = F, echo = F}

source("~/Documents/DSI/Burn/burn_data_trees.R")
```

## A. Accuracy of current criteria

```{r current_ghsirs, cache = T, echo = F, warning = F, message = F}


library(dplyr)
# Confusion matrix for different criteria 

# First blood infection day
TT.tree.data %>% filter(before_blood == TRUE | (first_blood_l == TRUE | n_blood == 0))  %>% 
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  select(Greenhalgh, first_blood_l) %>% table()

TT.tree.data %>% filter(before_blood == TRUE | (first_blood_l == TRUE | n_blood == 0))  %>% 
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  select(SIRS, first_blood_l) %>% table()

# Day before infection
TT.tree.data %>% filter(before_blood == TRUE | (blood_onset_tomorrow == TRUE | n_blood == 0))  %>% 
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  select(Greenhalgh, blood_onset_tomorrow) %>% table()

TT.tree.data %>% filter(before_blood == TRUE | (blood_onset_tomorrow == TRUE | n_blood == 0))  %>% 
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  select(SIRS, blood_onset_tomorrow) %>% table()

# Up to two days before (and including day of) infection
TT.tree.data %>% filter(before_blood == TRUE | (blood_onset_tomorrow == TRUE | n_blood == 0))  %>% 
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  mutate(days_before2 = (days_until_first_blood <=2 & days_until_first_blood >=0),
         days_before2 = replace(days_before2, is.na(days_before2), FALSE)) %>%
  select(Greenhalgh, days_before2) %>% 
  table()

TT.tree.data %>% filter(before_blood == TRUE | (blood_onset_tomorrow == TRUE | n_blood == 0))  %>%
  dplyr::filter(!is.na(V_TIME_PERFORMED_1) | !is.na(V_TIME_PERFORMED_2)) %>%
  mutate(days_before2 = (days_until_first_blood <=2 & days_until_first_blood >=0),
         days_before2 = replace(days_before2, is.na(days_before2), FALSE)) %>%
  select(SIRS, days_before2) %>% 
  table()

```

## B. Decision tree models

### Greenhalgh variables only 

```{r gh, cache = T, echo = F}


# Show tree plts
par(mfrow = c(1, 3))
gh_tree_first_blood$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)
gh_tree_day_before$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)

```

### SIRS variables only 

```{r sirs, cache = T, echo = F}


# Show tree plts
par(mfrow = c(1, 3))
sirs_tree_first_blood$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)
sirs_tree_day_before$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)

```


### Greenhalth and SIRS variables 

```{r ghsirs, cache = T, echo = F}


# Show tree plts
par(mfrow = c(1, 3))
ghsirs_tree_first_blood$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)
ghsirs_tree_day_before$pruned_tree %>% rpart.plot(., roundint = FALSE, digits = ndigits, extra = 1)

```

### Compare ROC plots

```{r compare_ROC_1, cache = T, echo = F, fig.height = 12, fig.width = 10}


par(mfrow = c(2,2))
plot_etc(sirs_tree_first_blood, sirs_tree_day_before, sirs_tree_multi_day, main = "SIRS variables only ")
plot_etc(gh_tree_first_blood, gh_tree_day_before, gh_tree_multi_day, main = "Greenhalgh variables only ")

```

### Compare confusion matrices


#### Greenhalgh variables
```{r compare_CM_Greenhalgh, cache = T, echo = c(1,4,6,9)}
# First blood infection day
cm1 = confusion_etc(gh_tree_first_blood$pruned_pred, seq(0,.6, .2))
cm1
# Ratios of false positive to true positives in above matrices:
sapply(cm1, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})
# Day before first blood infection day
cm2 = confusion_etc(gh_tree_day_before$pruned_pred, seq(0,.6, .2))
cm2
# Ratios of false positive to true positives in above matrices:
sapply(cm2, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})

```

####  SIRS variables
```{r compare_CM_SIRS, cache = T, echo = c(1,4,6,9)}
# First blood infection day
cm1 = confusion_etc(sirs_tree_first_blood$pruned_pred, seq(0,.6, .2))
cm1
# Ratios of false positive to true positives in above matrices:
sapply(cm1, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})
# Day before first blood infection day
cm2 = confusion_etc(sirs_tree_day_before$pruned_pred, seq(0,.6, .2))
cm2
# Ratios of false positive to true positives in above matrices:
sapply(cm2, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})

```

#### Greenhalgh and SIRS variables 
```{r compare_CM_ghsirs, cache = T, echo = c(1,4,6,9)}
# First blood infection day
cm1 = confusion_etc(ghsirs_tree_first_blood$pruned_pred, seq(0,.6, .2))
cm1
# Ratios of false positive to true positives in above matrices:
sapply(cm1, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})
# Day before first blood infection day
cm2 = confusion_etc(ghsirs_tree_day_before$pruned_pred, seq(0,.6, .2))
cm2
# Ratios of false positive to true positives in above matrices:
sapply(cm2, function(x) {x[nrow(x), 1]/ x[nrow(x), 2]})

```
